---
title: Practical Machine Learning Project Report
author: N. Kirnosov
output:
html_document:
fig_height: 10
fig_width: 10
---
        
## Introduction  
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible 
to collect a large amount of data about personal activity relatively inexpensively. 
These type of devices are part of the quantified self movement – a group of 
enthusiasts who take measurements about themselves regularly to improve their 
health, to find patterns in their behavior, or because they are tech geeks. 
One thing that people regularly do is quantify how much of a particular activity 
they do, but they rarely quantify how well they do it.  

In this project, we will use data from accelerometers on the belt, forearm, 
arm, and dumbell of 6 participants to predict the manner in which they 
did the exercise.  

## Libraries Used
```{r, cache = TRUE, message=FALSE, hide=TRUE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```
### Download the Data
```{r, cache = TRUE}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
        dir.create("./data")
}
if (!file.exists(trainFile)) {
        download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
        download.file(testUrl, destfile=testFile, method="curl")
}
```  
### Read the Data
After downloading the data from the data source, we can read the two csv files into two data frames.  
```{r, cache = T}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```
The training data set contains 19622 observations and 160 variables, while the testing data set contains 20 observations and 160 variables. The "classe" variable in the training set is the outcome to predict. 
